Edge sign prediction in general graphs and Correlation Clustering

### 2017-10-02 troll/supp_body.tex
add new dataset, competitor and results
only one section for troll additional material
explain better the cutsize with a figure

\section{Proofs from \autoref{s:algbatch}}

\subsection{Proof of \autoref{t:active}}
\label{ssec:troll_proof_bayes}

@@

\begin{proof}[Proof of \autoref{l:active}]
We apply \autoref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\log
\frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and, for all $u \in
V$, of $\Nin(u)$ and $\Nout(u)$. We have that, with probability at least $1-\delta$, at least $Q$
edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(v)$ are sampled for each $v$ such that
$\NNin(v) \ge \theta$, and at least $Q$ edges of $\Nout(u)$ are sampled for each $u$ such that
$\NNout(u) \ge \theta$.
For all $(u,v) \in E_{\theta}$ let
\[
  \pbar_v = \frac{1}{\din(v)}\sum_{u\in\NNin(v)} p_u \quad\text{and}\quad
  \qbar_u = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)} q_v
\]
and set for brevity $\hdeltain(v) = 1-\hun(v)$ and $\hdeltaout(u) = 1-\htr(u)$.
We now prove that $\hdeltain(v)$ and $\hdeltaout(u)$ are concentrated around their expectations for
all $(u,v) \in E_{\theta}$. Consider $\hdeltaout(u)$ (the same argument works for $\hdeltain(v)$).
Let $V_1,\dots,V_Q$ be the first $Q$ draws in $E_0 \cap \NNout(u)$ and define
\[
  \muhat_p(u) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_u+q_{V_t}}{2}~.
\]
Applying \autoref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(u)}$, and using
our choice of $Q$, we get that $\big|\muhat_p(u) - \mu_p(u)\big| \le \ve$ holds with probability at
least $1-\delta/(2|V|)$, where
\[
  \mu_p(u) = \frac{1}{\dout(u)}\sum_{v \in \NNout(u)} \frac{p_u + q_v}{2} = \frac{p_u + \qbar_u}{2}~.
\]
Now consider the random variables $Z_t = \Ind{y_{u,v_t}=1}$, for $t=1,\dots,Q$. Conditioned on
$V_1,\dots,V_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid V_t] =
\tfrac{p_u+q_{V_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and
using our choice of $Q$, we get that
\[
  \left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(u) \right| \le \ve 
\]
with probability at least $1-\delta/(2|V|)$ for every realization of $V_1,\dots,V_Q$.

@@

\subsection{Derivation of the maximum likelihood equations}
\label{ssec:troll_mle_deriv}

Recall that the training set $\trainset = \theset{\big(u_t,v_t),y_{u_t,v_t}\big)}{t=1,\dots,m}$
is drawn uniformly at random from $E$ without replacement. 
We can write

\begin{align*}
  \Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) = 
  \frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=+1}}\,
  \times\,\prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=-1}}\\
  &=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=+1}}\,\times\,
    \prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=-1}}\Biggl)
\end{align*}

so that $\log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)$ is proportional to

@@

\begin{equation*}
  \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
  = \sum_{k=1}^m 
  \frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
   - \,
  \sum_{k=1}^m \frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}}\,.
\end{equation*}

@@

\begin{align*}
  \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=+1}}\,
\times\,
    \prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=-1}}\Biggl)
\end{align*}

@@

\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}
= \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
- \, \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,.
\end{align*}

\subsection{Label propagation on $G''$}
\label{ssec:troll_lprop_weights}

@@

\begin{equation*}
  \frac{1}{16} \sum_{u,v \in E} \Bigl[ \eps\left(y_{u,v} - p_u'\right)^2 + \eps\left(y_{u,v} - q_v'\right)^2 + (p'_i+q_v')^2 \Bigr]
\end{equation*}

@@

\section{Proof from \autoref{s:algonline}}

@@

\section{Further Experimental Results}

@@

\begin{table}[bt]
  \centering
  \caption{Normalized logistic regression coefficients averaged over 12 runs (with one standard
  deviation) \label{tab:coeff}}
  \small
  \begin{tabular}{lrccc}
    \toprule
    & $\frac{|\trainset{}|}{|E|}$ &              $w_2'$ & $\tau'$ \\
    % \multirow{5}{*}{\aut{}} & $3\%$  &  $0.909 \pm 0.07$  &  $0.592 \pm 0.04$  \\
    %                         & $9\%$  &  $0.985 \pm 0.04$  &  $0.707 \pm 0.03$  \\
    \multirow{5}{*}{\aut{}} & $5\%$  &  $0.965 \pm 0.04$  &  $0.662 \pm 0.03$  \\
                            & $10\%$ &  $0.983 \pm 0.03$  &  $0.705 \pm 0.02$  \\
                            & $15\%$ &  $1.001 \pm 0.03$  &  $0.729 \pm 0.03$  \\
                            & $20\%$ &  $1.013 \pm 0.02$  &  $0.747 \pm 0.02$  \\
                            & $25\%$ &  $1.011 \pm 0.02$  &  $0.746 \pm 0.01$  \\
    \midrule
    % \multirow{5}{*}{\wik{}} & $3\%$  &  $0.919 \pm 0.04$  &  $0.666 \pm 0.02$  \\
    %                         & $9\%$  &  $0.940 \pm 0.03$  &  $0.726 \pm 0.01$  \\
    \multirow{5}{*}{\wik{}} & $5\%$  &  $0.920 \pm 0.02$  &  $0.691 \pm 0.02$  \\
                            & $10\%$ &  $0.940 \pm 0.01$  &  $0.730 \pm 0.01$  \\
                            & $15\%$ &  $0.947 \pm 0.01$  &  $0.741 \pm 0.01$  \\
                            & $20\%$ &  $0.963 \pm 0.01$  &  $0.760 \pm 0.01$  \\
                            & $25\%$ &  $0.962 \pm 0.02$  &  $0.764 \pm 0.01$  \\
    \midrule
    % \multirow{5}{*}{\sla{}} & $3\%$  &  $1.012 \pm 0.02$  &  $0.669 \pm 0.01$  \\
    %                         & $9\%$  &  $1.014 \pm 0.01$  &  $0.703 \pm 0.01$  \\
    \multirow{5}{*}{\sla{}} & $5\%$  &  $1.024 \pm 0.02$  &  $0.693 \pm 0.01$  \\
                            & $10\%$ &  $1.017 \pm 0.01$  &  $0.705 \pm 0.01$  \\
                            & $15\%$ &  $1.007 \pm 0.01$  &  $0.707 \pm 0.01$  \\
                            & $20\%$ &  $1.002 \pm 0.00$  &  $0.710 \pm 0.00$  \\
                            & $25\%$ &  $0.995 \pm 0.01$  &  $0.712 \pm 0.00$  \\
    \midrule
    % \multirow{5}{*}{\epi{}} & $3\%$  &  $1.124 \pm 0.02$  &  $0.786 \pm 0.02$  \\
    %                         & $9\%$  &  $1.066 \pm 0.01$  &  $0.783 \pm 0.01$  \\
    \multirow{5}{*}{\epi{}} & $5\%$  &  $1.099 \pm 0.02$  &  $0.791 \pm 0.02$  \\
                            & $10\%$ &  $1.059 \pm 0.01$  &  $0.782 \pm 0.01$  \\
                            & $15\%$ &  $1.037 \pm 0.01$  &  $0.774 \pm 0.01$  \\
                            & $20\%$ &  $1.018 \pm 0.01$  &  $0.765 \pm 0.01$  \\
                            & $25\%$ &  $1.010 \pm 0.01$  &  $0.763 \pm 0.01$  \\
    \midrule
    % \multirow{5}{*}{\kiw{}} & $3\%$  &  $1.075 \pm 0.02$  &  $0.851 \pm 0.02$  \\
    %                         & $9\%$  &  $1.041 \pm 0.01$  &  $0.872 \pm 0.01$  \\
    \multirow{5}{*}{\kiw{}} & $5\%$  &  $1.047 \pm 0.02$  &  $0.853 \pm 0.01$  \\
                            & $10\%$ &  $1.038 \pm 0.01$  &  $0.872 \pm 0.01$  \\
                            & $15\%$ &  $1.025 \pm 0.01$  &  $0.876 \pm 0.01$  \\
                            & $20\%$ &  $1.012 \pm 0.01$  &  $0.874 \pm 0.01$  \\
                            & $25\%$ &  $1.007 \pm 0.01$  &  $0.874 \pm 0.01$  \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

% \section{Proofs from \autoref{s:algbatch}}
\section{Additional material}
\subsection{Proofs from \autoref{s:algbatch}}

\subsubsection{Proof of \autoref{t:active}}
\label{ssec:troll_proof_bayes}

@@

\begin{proof}[Proof of \autoref{l:active}]
We apply \autoref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\log
\frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and, for all $u \in
V$, of $\Nin(u)$ and $\Nout(u)$. We have that, with probability at least $1-\delta$, at least $Q$
edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(v)$ are sampled for each $v$ such that
$\NNin(v) \ge \theta$, and at least $Q$ edges of $\Nout(u)$ are sampled for each $u$ such that
$\NNout(u) \ge \theta$.
For all $(u,v) \in E_{\theta}$ let
\[
  \pbar_v = \frac{1}{\din(v)}\sum_{u\in\NNin(v)} p_u \quad\text{and}\quad
  \qbar_u = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)} q_v
\]
and set for brevity $\hdeltain(v) = 1-\hun(v)$ and $\hdeltaout(u) = 1-\htr(u)$.
We now prove that $\hdeltain(v)$ and $\hdeltaout(u)$ are concentrated around their expectations for
all $(u,v) \in E_{\theta}$. Consider $\hdeltaout(u)$ (the same argument works for $\hdeltain(v)$).
Let $V_1,\dots,V_Q$ be the first $Q$ draws in $E_0 \cap \NNout(u)$ and define
\[
  \muhat_p(u) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_u+q_{V_t}}{2}~.
\]
Applying \autoref{l:hoeff-wr} to the set $\theset{\tfrac{p_u+q_v}{2}}{v \in \NNout(u)}$, and using
our choice of $Q$, we get that $\big|\muhat_p(u) - \mu_p(u)\big| \le \ve$ holds with probability at
least $1-\delta/(2|V|)$, where
\[
  \mu_p(u) = \frac{1}{\dout(u)}\sum_{v \in \NNout(u)} \frac{p_u + q_v}{2} = \frac{p_u + \qbar_u}{2}~.
\]
Now consider the random variables $Z_t = \Ind{y_{u,v_t}=1}$, for $t=1,\dots,Q$. Conditioned on
$V_1,\dots,V_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid V_t] =
\tfrac{p_u+q_{V_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and
using our choice of $Q$, we get that
\[
  \left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(u) \right| \le \ve 
\]
with probability at least $1-\delta/(2|V|)$ for every realization of $V_1,\dots,V_Q$.

@@

\subsubsection{Derivation of the maximum likelihood equations}
\label{ssec:troll_mle_deriv}

Recall that the training set $\trainset = \theset{\big(u_k,v_k),y_{u_k,v_k}\big)}{k=1,\dots,m}$
is drawn uniformly at random from $E$ without replacement. 
We can write

\begin{align*}
  \Pr&\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right) = 
  \frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=+1}}\,
  \times\,\prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=-1}}\\
  &=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=+1}}\,\times\,
    \prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=-1}}\Biggl)
\end{align*}

so that $\log \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)$ is proportional to

@@

\begin{equation*}
  \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)}{\partial p_{\ell}}
  = \sum_{k=1}^m 
  \frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
   - \,
  \sum_{k=1}^m \frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}}\,.
\end{equation*}

@@

\begin{align*}
  \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right) 
=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=+1}}\,
\times\,
    \prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=-1}}\Biggl)
\end{align*}

@@

\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)}{\partial q_{\ell}}
= \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
- \, \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,.
\end{align*}

\subsubsection{Label propagation on $G''$}
\label{ssec:troll_lprop_weights}

@@

\begin{equation*}
  \frac{1}{16} \sum_{u,v \in E} \Bigl[ \eps\left(y_{u,v} - p_u'\right)^2 + \eps\left(y_{u,v} - q_v'\right)^2 + (p'_u+q_v')^2 \Bigr]
\end{equation*}

@@

\subsection{Proof from \autoref{s:algonline}}

@@

\subsection{Further Experimental Results}

@@

\input{all_coeff}
